`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys

### 2019-05-31 analysis/paper/sim_survey_paper.Rmd
Simplify the introduction

- As with any population monitoring program, accurate and reliable results from a fisheries-independent trawl survey depends on good survey design. Ideally, the sampling design of large and costly surveys should be simulation tested before embarking on an expensive field program.

- While there are several simulation frameworks for testing the design of fisheries-independent trawl surveys [e.g. @puerta2019; @schnute2003], there are few tools that facilitate such work. Here we document an R package with a series of functions for testing the multi-layered sampling design of spatial, age-structured populations. The package can also serve as a medium for the simulation testing of methods for analyzing such data.

It is well known that the accuracy and reliability of any population monitoring program hinges on good survey design and analysis. Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on expensive field programs. This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numurous regional fisheries management organizations. These surveys have become an increasingly important research tool for the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. While costly to obtain, this information forms the basis of of many stock assessments used throughout the world. Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection. Solving this optimization problem is not a trivial task given the complexity of the population processes we aim to represent and the multi-stage nature of the typical sampling program.

The complexity of the system and the surveys may be why simulation tests of the full sampling and analylicical pathway are rare. Simulations provide a playground of sorts...multiple frameworks...few accessible...

Fisheries-independent trawl surveys have become an increasingly important research tool for the management of dynamic fish stocks. These surveys provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. While costly to obtain, this information forms the basis of of many stock assessments models used throughout the world. Effective and informed management decisions therefore require surveys that maximize information while minimizing the expense of data collection. Determining the number of samples required to adequately characterize a fish population is particularly challenging since data tend to be collected across one or more stages [@Aanes2015]. Optimal survey design is further complicated by the fact that trawls sample clusters of fish that tend to have similar characteristics, such as length and age. This phenomenon results in samples with positive intra-haul correlation, which can markedly reduce the effective sample size for estimating length and age frequencies of the target population [@Pennington1994; @Pennington2002; @Stewart2014]. The complexities of multi-stage sampling and intra-haul correlation make it difficult to answer the question *"how much sampling is enough?"* Moreover, the answer to this question is inextricably linked with the way in witch these data are analyzed.

While analyses of fisheries-independent survey data has typically been confined to design-based approaches [e.g. stratified-random survey data tend to be analyzed using standard stratified-analysis formulae;  @Cochran1977], there are a growing number of model-based approaches for analyzing such data [e.g. @smith1990; @Berg2014; @Thorson2015]. The merits of both approaches have long been debated, however, simulation tests of the full sampling and analylitical pathway are rarely conducted. This is especially true for design- or model-based estimates of abundance-at-age, which is a criticial input to any age-based stock assessment model. Ideally, the design and analysis of any large and costly survey should be simulation tested before embarking on an expensive field program.

Two main approaches have been applied to try and answer this question: one involves statistical estimators of effective sample size [@Pennington2002; @Stewart2014], and the other employs resampling methods to test alternate sampling protocol [@Cervino2006; @Zhang2013]. Both approaches have shown that lengths are usually oversampled and a reduction in said sampling can often be done without significant loss of accuracy [@Pennington2002]. The same result holds for other positively correlated characteristics, such as age [@Coggins2013] and stomach contents [@Bogstad1995]. Pennington and Vølstad [-@Pennington1994] concluded that the best way to improve the accuracy of surveys with strong intra-haul correlation is to take fewer samples per sampling unit and increase the total number of sampling locations. Few studies, however, have tested these conclusions using simulations. Simulations are a useful approach for informing the design of surveys [@sutherland2006] and they have been used widely for exploring the number of samples needed to obtain estimates with reasonable precision [e.g. distance sampling surveys; @thomas2010]. A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled (set, length and age sampling protocol). The relative effects of each stage of sampling on our perception of the "truth" can then be tested. This level of testing is of particular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling. A key challenge, of course, is building a simulation that emulates reality.

The objective of this paper is to use a realistic simulation to evaluate the efficacy of various sampling protocol. First we simulated a spatially autocorrelated target population, and then we simulated surveys over this field. Set density was varied across surveys, as was length and age sampling efforts. Estimates of abundance at age were then obtained using a standard stratified approach [@Smith1981] and deviation from the truth was assessed. Given previous findings, we predict that the greatest gains in precision will come from increasing the number of hauls; increased sampling of correlated variables (length and age) on a set by set basis will be relatively ineffective. The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis. While our case study is specific, we also hope that the approach presented here will provide a framework for optimizing sampling protocol for an array of species with different life-histories and distributions. To facilitate future work and collaborations, we have developed an R package for conducting the tests presented in this paper (https://github.com/PaulRegular/SimSurvey).

@@

The **`SimSurvey`** package was written in the programming language R [@R] and, in short, the package allows for the simulation of stratified-random surveys of an age-structured population that varies in space and time. **`SimSurvey`** relies heavily on functions from the data.table [@dowle2017], raster [@hijmans2016] and plotly [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. All the source R code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the 0.1.0 release of the package can be installed via GitHub:

```{r, echo = TRUE, eval = FALSE}
install.packages("devtools")
devtools::install_github("PaulRegular/SimSurvey@v0.1.0") 
```

@@

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e. $R^2$), and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate abundance] section. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division.

@@

The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs [e.g. @Cochran1977]. Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates [e.g. @smith1990; @Berg2014; @Thorson2015]. **`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.). Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested. Existing and future approaches for calculating age-based indices of abundance can be simulation tested using *`SimSurvey`*.

>>>>

It is well known that the accuracy and reliability of results from any population monitoring program hinges on good survey design and analysis [@thomas2010]. Ideally, the design and analysis of large and costly survey should be simulation tested before embarking on expensive field programs. This is especially true for the large-scale and costly fisheries-independent trawl surveys conducted by numerous regional fisheries management organizations. These surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. This information forms the basis of of many stock assessments throughout the world. Efficient, effective and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection. While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent. The complexity of this problem is perhaps why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].

Here we document **`SimSurvey`**, an **`R`** package designed to facilitate realistic simulations of fisheries-independent trawl surveys. This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused here on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below, with specific emphasis of a case study that motivated this work.

@@

The **`SimSurvey`** package was written in the programming language **`R`** [@R] and, in short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time. **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. All the source **`R`** code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey) and the package can be installed via GitHub:

```{r, echo = TRUE, eval = FALSE}
install.packages("devtools")
devtools::install_github("PaulRegular/SimSurvey")
```

@@

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell (i.e. $R^2$), and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate abundance] section. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division. Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish sampled. Age determination is especially time-consuming, which is why age sampling tends to be length-stratified.

@@

The analysis of data from fisheries-independent surveys have generally been confined to design-based mean and variance estimates of abundance using standard formula for stratified-random designs [e.g. @Cochran1977]. Nevertheless, there has long been interest in using model-based approaches to improve abundance estimates [e.g. @smith1990; @Berg2014; @Thorson2015]. **`SimSurvey`** can serve as a convenient tool for simulation testing mean and variance estimates provided by a range of different approaches (design-based analyses, bootstrap estimates, generalized additive models, geostatistical models, etc.). Moreover, the full analytical pathway for obtaining age-disaggregated estimates of abundance has rarely been simulation tested. Existing and future approaches for calculating age-based indices of abundance can be simulation tested using **`SimSurvey`**.
